Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-09-21 grove-paper.tex
Rough text for Section 2.3.5 Single-User Cycles (Michael D. Adams)

Note that this edge has a new identifier distinct from the one before.
In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
Instead a new edge with a new identifiers is created.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
    {}
   },
   {}
  }
 },
 v2 -> v8
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Single-User Cycles}
\label{fig:Single-User Cycles}
\end{figure}

@@

Although this case is likely to be infrequent.

Since we are representing programs as graphs,
we have to consider cycles.

Cycles can occur in a both single-user and multi-user contexts.

In a single user context, a cycle can appear when a user pastes
a vertex as a child of itself.
An example of this is shown in \autoref{fig:Single-User Cycles},
where the user has pasted vertex~TODO as a child of vertex~TODO.

As a user interface consideration, it might be best to either
disallow such edits to at least warn users that their
edits would create a cycle.

@@

>>>>

Note that this edge has a new identifier distinct from the one before.
In our model, once an edge with a particular identifier is deleted, it cannot be recreated.
Instead a new edge with a new identifier is created.

@@

\begin{figure*}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
    {}
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Single-User Cycles-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
    {}
   },
   "$\texttt{+}_{18}$" [> "right", > "7"'] -> {
    "$\texttt{*}_{10}$" [> "left", > "19"'] -> {
     "$\texttt{var(x)}_{12}$" [> "left", > "19"'],
     {}
    },
    {}
   }
  }
 },
};
\end{tikzpicture}
\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by copy as the right child of \Vertex{TODO}}
\label{fig:Single-User Cycles-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   v2/"$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
    {}
   },
   {}
  }
 },
 v2 -> v8
};
\end{tikzpicture}
\caption{\autoref{fig:Single-User Cycles-a} after Alice copy and pastes \Vertex{TODO} by reference as the right child of \Vertex{TODO}}
\label{fig:Single-User Cycles-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Single-User Cycles}
\label{fig:Single-User Cycles}
\end{figure*}

@@

Another case we must consider is when cycles appear in the graph.
We categorize these into cycles caused by the action of a single user
and cycles caused by the interaction of the actions of multiple users.

In a single user context,
normal insertion and deletion of code by the user cannot create cycles.
However, it is possible with certain kinds of copy and paste.
For example, suppose Alice is editing the code in \autoref{fig:Single-User Cycles-a}
and uses copy and paste to copy \Vertex{TODO} to the right child of \Vertex{TODO}.
There are two ways to interpret the paste action.
The first interpretation is to create a deep copy of \Vertex{TODO}.
This results in \autoref{fig:Single-User Cycles-b} and
does not cause a cycle.
The second interpretation is to simply add an edge to \Vertex{TODO}.
This results in \autoref{fig:Single-User Cycles-b}
and causes a cycle.

Note that not all pastes should be deep copies.
For example, Alice may have accomplished code move in \autoref{sub:Editing Nested Parts of the Code}
done by a cutting from the old position and pasting to the new position.
Preserving Bob's nested edits requires that the paste be by reference instead of by copy.
Distinguishing when a paste should be by reference versus by copy
is ultimately a user interface question.
Cycles caused by the local user's edits be detected as soon as a user enters them
by noting either that the graph would contain a cycle or the vertex
already has a parent somewhere in the graph.

Thus, as a user interface consideration, it might be best to either
disallow such edits to at least warn users that their
edits would create a cycle.

@@

\subsection{Summary}
